---
title: "MLR_Quiz-1"
author: "Darrin"
date: "2025-01-26"
output: word_document
---

```{r}
library(tidyverse) #tidyverse set of packages and functions
library(tidymodels)
library(glmnet) #for Lasso, ridge, and elastic net models 
library(GGally) #create ggcorr and ggpairs plots
library(ggcorrplot) #create an alternative to ggcorr plots
library(MASS) #access to forward and backward selection algorithms
library(car) #best subset selection
library(lubridate) 
library(lmtest) #for the dw test
library(splines) #for nonlinear fitting
```

As we did before, read in the data and select a few variables of interest (to keep the analysis from being too large).     
```{r}
bike = read_csv("bike_cleaned-3.csv")
```

Examine the bike data frame.  
```{r}
str(bike)
summary(bike)
```


Convert “dteday” from a character variable to a date variable. The code below will perform this conversion:
bike = bike %>% mutate(dteday = mdy(dteday))
#Note that mdy is a lubridate package function
#You can read more about lubridate here: https://lubridate.tidyverse.org/
```{r}
bike = bike %>% mutate(dteday = mdy(dteday))
```


Convert the remaining character variables to factors. You can do this one variable at a time or use a
“mutate_if”. This function examines each variable. If the variable is a character it is converted into a factor.
Otherwise, the variable is left alone.
bike = bike %>% mutate_if(is.character, as_factor)
```{r}
bike = bike %>% mutate_if(is.character, as_factor)
```

Finally, convert the “hr” variable into a factor. We do because, even though “hr” is numeric, we want to try
each hour as a category. This can be a useful trick when you have a numeric variable with only a few unique
values (DO NOT do this for numeric variables that are continuous and contain many unique values) and
when the relationship between the numeric variable and the response variable is clearly nonlinear (as we will
see in a moment when we plot “hr” versus the response variable).
bike = bike %>% mutate(hr = as_factor(hr))
```{r}
bike = bike %>% mutate(hr = as_factor(hr))
```


```{r}
str(bike)
summary(bike)
```

Next we look at correlation. This is a logical step since almost all of our variables are quantitative.    
```{r}
#use "ggcorr" to create a correlation matrix with labels and correlation reported to two decimals
ggcorr(bike, label = "TRUE", label_round = 2) 

```

Question 1 Which of the quantitative variables appears to be best correlated with “count”? NOTE: Ignore
the “registered” and “casual” variable as the sum of these two variables equals “count”. Because these
variables combine to make the response variable, they cannot be used as predictors. You can also ignore the
“instant” variable as it is just a row number.  D. temp
A. windspeed
B. hum
C. atemp
D. temp
——
Correlation and Categorical Variables We cannot use correlation to assess the relationship between
a categorical predictor variable and our response variable. A good option is to visualize the relationship
between the categorical and response variables via a boxplot (other visualizations can work too, but a boxplot
is often a good place to start). Note that the categorical variable should be on the x-axis.

```{r}
ggplot(bike,aes(x=hr,y=count)) + geom_boxplot() + geom_jitter(alpha = 0.15) + theme_bw()
```
If you create a boxplot for “hr” and “count” you will see that it is fairly obvious that “hr” affects “count”. It
should also be obvious that the relationship between “hr” and “count” is not linear.

Repeat this boxplot-based analysis for each of the categorical variables.


```{r}
ggplot(bike,aes(x=season,y=count)) + geom_boxplot() + geom_jitter(alpha = 0.15) + theme_bw()
```
Question 2 Which “season” appears to have the highest count of rides?  C. Summer

A. Winter
B. Spring
C. Summer
D. Fall


### Models
The first model we'll build uses the variable that is best correlated with "count", "hr". This is a univariate (simple) linear regression model. We also plot this model.  
```{r}
bike_recipe = recipe(count ~ hr, bike)

lm_model = #give the model type a name 
  linear_reg() %>% #specify that we are doing linear regression
  set_engine("lm") #specify the specify type of linear tool we want to use 

lm_wflow = 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(bike_recipe)

lm_fit = fit(lm_wflow, bike)
```

```{r}
summary(lm_fit$fit$fit$fit)
```

```{r}
ggplot(bike, aes(x=hr, y=count)) + geom_point() + geom_smooth(method = lm, se = FALSE) + theme_bw()
```
Question 3 Build a linear regression model (using tidymodels) with “hr” to predict “count”. You will use
this model to answer the next several questions.
How many dummy (indicator) variables are used to represent “hr” in the model? 23

Question 4 In your model from Question 3, which hour is selected as the “base” level (category)? The base
level does not have an associated coefficient (slope) in the linear regression model. 0

Question 5 During which hour of the day does the model predict the highest number of rides? 17

```{r}
ggplot(bike, aes(x=temp, y=count)) + geom_point() + geom_smooth(method = lm, se = FALSE) + theme_bw()
```
Question 6 Plot “temp” (x axis) versus “count” (y axis) using an appropriate plot type.
Which statement best describes the general relationship between “temp” and “count”? A. As “temp” increases, “count” appears to generally increase.
A. As “temp” increases, “count” appears to generally increase.
B. As “temp” increases, “count” appears to generally decrease.
C. There does not appear to be a relationship between “temp” and “count”

```{r}
bike_recipe = recipe(count ~ hr + temp, bike)

lm_model = #give the model type a name 
  linear_reg() %>% #specify that we are doing linear regression
  set_engine("lm") #specify the specify type of linear tool we want to use 

lm_wflow = 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(bike_recipe)

lm_fit = fit(lm_wflow, bike)
```

```{r}
summary(lm_fit$fit$fit$fit)
```
Question 7 Create a linear regression model (using tidymodels) with “hr” and “temp” to predict “count”.
You will use this model to answer the next several questions. 
What is the value of the slope coefficient for “hr23” in this model (to three decimal places)? 31.748

Question 8 What is the adjusted R-squared value (to four decimal places) for the model from Question 7?  0.5886

```{r}
bike_recipe = recipe(count ~ temp + atemp, bike)

lm_model = #give the model type a name 
  linear_reg() %>% #specify that we are doing linear regression
  set_engine("lm") #specify the specify type of linear tool we want to use 

lm_wflow = 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(bike_recipe)

lm_fit = fit(lm_wflow, bike)
```

```{r}
summary(lm_fit$fit$fit$fit)
```

Question 9 Create a linear regression model (using tidymodels as usual) with “temp” and “atemp” to predict
“count”. What is the adjusted R-squared value (to four decimal places) of this model? 0.1638

Question 10 Which of the two variables in the model from Question 9 are significant? A. temp Only
A. temp ONLY
B. atemp ONLY
C. Neither temp nor atemp are significant
D. Both temp and atemp are significant

Question 11 The model from Question 9 likely demonstrates which phenomenon?  C. Multicollinearity
A. Non-constant variance of residuals
B. Non-normality of residuals
C. Multicollinearity
D. None of these

```{r}
bike2 = bike %>% dplyr::select("season", "mnth", "hr", "holiday", "weekday", "workingday", "weathersit", "temp", "atemp", "hum", "windspeed")
summary(bike2)
```
```{R}
allmod = lm(count ~., bike) #use the ~. to include all predictors rather than typing them all in
summary(allmod)
```
Question 12 Build a backward stepwise regression model to predict “count”. Your “allmod” (the starting
model) should include the following variables: season, mnth, hr, holiday, weekday, workingday, weathersit,
temp, atemp, hum, and windspeed.
In the “allmod” you should see that the “workingday” variable appears with “NA” values in the model
summary. This is happening because “workingday” is a perfect combination of two other predictor variables.
Which two variables combine to make “workingday”? B. weekday and holiday
A. season and mnth
B. weekday and holiday
C. hr and mnth
D. season and mnth
Backward stepwise  
```{r}
#backward
backmod = stepAIC(allmod, direction = "backward", trace = TRUE) #trace = TRUE shows how the model is built (which variables are removed)
summary(backmod)
```
Question 13 The backward stepwise method removes only one variable. Which variable is removed?  B. workingday
A. windspeed
B. workingday
C. hum
D. holiday
