---
title: "Mod2 Simple Linear Regression Quiz"
author: "Darrin"
date: "2025-01-25"
output: word_document
---

Libraries
```{r,warning=FALSE,message=FALSE}
library(tidyverse)
library(tidymodels)
library(GGally)
library(lmtest) #for the Durbin-Watson test
```

Read-in the dataset
```{R}
air = airquality
```

Details concerning this dataset can be found here: http://rpubs.com/Nitika/linearRegression_Airquality.

Examine the structure and summary of the dataset  
```{r}
str(air) #all variables numeric
summary(air) #no missingness
```
Question 1 How many rows are in the “air” dataframe?  153
Question 2 How many columns are in the “air” dataframe? 6
Question 3 True/False: There is missing data in “Ozone” variable in the dataframe. True

Let's look at ggpairs plot for visualization and correlation.  
```{r}
ggpairs(air)
```
Question 4 Which variable is most likely to be the response (Y) variable? D. Temp
A. Ozone
B. Solar.R
C. Wind
D. Temp
E. Month
F. Day


We have three approaches that we can typically select from to deal with missing data:
1. Delete the rows with missing data
2. Delete the columns with missing data
3. Impute (i.e., estimate or guess) values to replace the missing values

Here we’ll choose to delete rows with any missing data. Use the code below to apply the “drop_na” function
to the “air” dataframe. The resulting dataframe will be called “air2”. You will use this dataframe for the
remainder of the assignment.
air2 = air %>% drop_na()
```{r}
air2 = air %>% drop_na()
```
Examine the structure and summary of the dataset  
```{r}
str(air2) #all variables numeric
summary(air2) #no missingness
```

Question 5 How many rows remain in this new (air2) data frame? 111
Question 6 How many columns remain in this new (air2) data frame?  6

Use the “ggpairs” function to develop a visualization of the relationships in this dataset and to show correlation
values for the combinations of variables.
Let's look at ggpairs plot for visualization and correlation.  
```{r}
ggpairs(air2)
```
Then use the “ggcorr” function to develop a correlation matrix for the variables. Hint: Use “label = TRUE”
in the “ggcorr” function to show the correlation values.
```{r}
ggcorr(air2, label = TRUE)
```

Question 7 Which variable is most strongly correlated with the “Ozone” variable? C. Temp
1
A. Solar.R
B. Wind
C. Temp
D. Month
E. Day
Question 8 Which variable is least strongly correlated with the “Ozone” variable? E. Day
A. Solar.R
B. Wind
C. Temp
D. Month
E. Day

```{r}
ggplot(air2, aes(x=Temp,y=Ozone)) + geom_point(alpha=0.1) + geom_smooth(method = "lm", color = "red") + theme_bw()
```

Question 9 Plot “Temp” (x axis) versus “Ozone” (y axis) using the “ggplot” function. Choose an appropriate
chart type. Which statement best describes the relationship between “Temp” and “Ozone”? C. As Temp increases, Ozone increases
A. As Temp increases, Ozone decreases
B. As Temp increases there is no noticeable change in Ozone
C. As Temp increases, Ozone increases


Use Tidymodels to create a linear regression model using “Temp” to predict “Ozone”. You miss wish to call
your model fit “lm_fit”.
```{r}
air_simple = recipe(Ozone ~ Temp, air2)
air_simple
```
Next we specify the type of model that we are building.  
```{r}
lm_model = #give the model type a name 
  linear_reg() %>% #specify that we are doing linear regression
  set_engine("lm") #specify the specify type of linear tool we want to use 
```

Next we combine the recipe and the model with a workflow.  
```{r}
lm_wflow = 
  workflow() %>% 
  add_model(lm_model) %>% 
  add_recipe(air_simple)
```

Next we fit (execute) the workflow on our dataset.  
```{r}
lm_fit = fit(lm_wflow, air2)
```

```{r}
summary(lm_fit$fit$fit$fit) #three fits :), the actual fit is embedded deeply in the object
```
Question 10 What is the slope of this regression model (to four decimal places)? 2.4391
Question 11 what is the R-squared value of this model (not Adjusted R-squared) (to three decimal places)? .488
Question 12 Is the “Temp” variables significant in the model? Yes


Question 13 Use the code below to generate 95% confidence intervals for the coefficients. Note that you
may need to change “lm_fit” to the name of your model fit if you used a different name.
True/False: A 95% confidence interval for the slope coefficient does not contain zero. True
confint(lm_fit$fit$fit$fit)
```{r}
confint(lm_fit$fit$fit$fit)
```


Question 14: Using your linear regression model with “Temp” to predict “Ozone”, what is the predicted
“Ozone” value when “Temp” is equal to 80 (to two decimal places)? 47.48
#Using predict function
```{r}
testdata = data.frame(Temp = c(80))
predict(lm_fit, new_data = testdata)
```

Question 15 Perform appropriate model diagnostics to verify whether or not the model appears to meet the
four linear regression model assumptions.
Let's look at this regression line on our plot.
```{r}
ggplot(air2, aes(x=Temp,y=Ozone)) + geom_point(alpha=0.1) + geom_smooth(method = "lm", color = "red") + theme_bw()
```
**Assumption 1:** The predictor and response variable have a linear relationship  
As noted above, it seems "reasonable" to say that there is something of a linear relationship between these two variables, but there is definitely a nonlinear effect present.  

**Assumption 2:** Model errors (residuals) are independent  
Let's use the Durbin-Watson Test to examine independence of residuals. The dwtest function is from the lmtest package.  
```{r}
dwtest(lm_fit$fit$fit$fit)
```
**Assumption 3:** Model residuals exhibit constant variance  
Examine a plot of residuals.  
```{r}
air3 = air2 %>% mutate(resid1 = lm_fit$fit$fit$fit$residuals) #add the model residuals to our data frame
ggplot(air3,aes(x=Temp,y=resid1)) + geom_point() + theme_bw()
```
**Assumption 4:** Model residuals are Normally-distributed  
Examine a histogram of the residuals.  
```{r}
ggplot(air3,aes(x=resid1)) + geom_histogram() + theme_bw()
```
True/False: There is no evidence of non-independent (autocorrelated) residuals.  True