---
title: "Project Phase 1"
author: "Darrin Pohlman"
date: "2025-02-15"
output: word_document
---

## Objective:  Identify which variables may be strong predictors of the "failure" variable and any other interesting descriptive findings

Libraries  
```{r}
library(tidyverse)
library(tidymodels)
library(VIM) #visualizing missingness
library(skimr)
library(GGally)
library(gridExtra)
# library(mice) #package for imputation
# library(glmnet) #for Lasso, ridge, and elastic net models 
```

Load data from the train.csv file.  
```{r}
train = read_csv("train.csv")
```

Structure and summary
```{r}
str(train)
summary(train)
```

Factor conversion. Convert and recode the response variable failure.
```{r}
train = train %>% mutate(failure = as_factor(failure)) %>% 
  mutate(failure = fct_recode(failure, "No" = "0", "Yes" = "1" )) 
```

Factor conversion attributes. Convert and recode the attribute variable.
```{r}
train = train %>% mutate(attribute_0 = as_factor(attribute_0)) %>% 
              mutate(attribute_0 = fct_recode(attribute_0, "7" = "material_7", "5" = "material_5")) 
train = train %>% mutate(attribute_1 = as_factor(attribute_1)) %>% 
              mutate(attribute_1 = fct_recode(attribute_1, "8" = "material_8", "5" = "material_5", "6" = "material_6")) 
```

Convert all character variables to factors  
```{r}
train = train %>% mutate_if(is.character,as_factor)
str(train)
summary(train)
```

```{r}
p1 = ggplot(train, aes(x = product_code, fill = failure)) + geom_bar(position = "fill") 
l2 = ggplot(train, aes(x = failure, y = loading)) + geom_boxplot()
grid.arrange(p1,l2, ncol = 2)
```


```{r}
#m1 = ggplot(train, aes(x = failure, y = measurement_0)) + geom_boxplot()
#m2 = ggplot(train, aes(x = failure, y = measurement_1)) + geom_boxplot()
#m3 = ggplot(train, aes(x = failure, y = measurement_2)) + geom_boxplot()
#m4 = ggplot(train, aes(x = failure, y = measurement_3)) + geom_boxplot()
#m5 = ggplot(train, aes(x = failure, y = measurement_4)) + geom_boxplot()
#m6 = ggplot(train, aes(x = failure, y = measurement_5)) + geom_boxplot()
#m7 = ggplot(train, aes(x = failure, y = measurement_6)) + geom_boxplot()
#m8 = ggplot(train, aes(x = failure, y = measurement_7)) + geom_boxplot()
#m9 = ggplot(train, aes(x = failure, y = measurement_8)) + geom_boxplot()
#m10 = ggplot(train, aes(x = failure, y = measurement_9)) + geom_boxplot()
#m11 = ggplot(train, aes(x = failure, y = measurement_10)) + geom_boxplot()
#m12 = ggplot(train, aes(x = failure, y = measurement_11)) + geom_boxplot()
#m13 = ggplot(train, aes(x = failure, y = measurement_12)) + geom_boxplot()
#m14 = ggplot(train, aes(x = failure, y = measurement_13)) + geom_boxplot()
#m15 = ggplot(train, aes(x = failure, y = measurement_14)) + geom_boxplot()
#m16 = ggplot(train, aes(x = failure, y = measurement_15)) + geom_boxplot()
#m17 = ggplot(train, aes(x = failure, y = measurement_16)) + geom_boxplot()
#m18 = ggplot(train, aes(x = failure, y = measurement_17)) + geom_boxplot()
#grid.arrange(m1,m2,m3,m4,m5,m6,m7,m8,m9,m10,m11,m12,m13,m14,m15,m16,m17,m18, ncol = 3)
```

```{r}
m1 = ggplot(train, aes(x = failure, y = measurement_0)) + geom_boxplot()
m2 = ggplot(train, aes(x = failure, y = measurement_1)) + geom_boxplot()
m3 = ggplot(train, aes(x = failure, y = measurement_2)) + geom_boxplot()
m4 = ggplot(train, aes(x = failure, y = measurement_3)) + geom_boxplot()
m5 = ggplot(train, aes(x = failure, y = measurement_4)) + geom_boxplot()
m6 = ggplot(train, aes(x = failure, y = measurement_5)) + geom_boxplot()

grid.arrange(m1,m2,m3,m4,m5,m6, ncol = 3)
```
```{r}
m7 = ggplot(train, aes(x = failure, y = measurement_6)) + geom_boxplot()
m8 = ggplot(train, aes(x = failure, y = measurement_7)) + geom_boxplot()
m9 = ggplot(train, aes(x = failure, y = measurement_8)) + geom_boxplot()
m10 = ggplot(train, aes(x = failure, y = measurement_9)) + geom_boxplot()
m11 = ggplot(train, aes(x = failure, y = measurement_10)) + geom_boxplot()
m12 = ggplot(train, aes(x = failure, y = measurement_11)) + geom_boxplot()
grid.arrange(m7,m8,m9,m10,m11,m12, ncol = 3)
```

```{r}
m13 = ggplot(train, aes(x = failure, y = measurement_12)) + geom_boxplot()
m14 = ggplot(train, aes(x = failure, y = measurement_13)) + geom_boxplot()
m15 = ggplot(train, aes(x = failure, y = measurement_14)) + geom_boxplot()
m16 = ggplot(train, aes(x = failure, y = measurement_15)) + geom_boxplot()
m17 = ggplot(train, aes(x = failure, y = measurement_16)) + geom_boxplot()
m18 = ggplot(train, aes(x = failure, y = measurement_17)) + geom_boxplot()
grid.arrange(m13,m14,m15,m16,m17,m18, ncol = 3)
```

Next we look at correlation. This is a logical step since almost all of our variables are quantitative.    
```{r}
#use "ggcorr" to create a correlation matrix with labels and correlation reported to two decimals
ggcorr(train, method=c("pairwise"),label = TRUE,label_color = "blue", label_round = 2, label_size = 2, hjust = 1, layout.exp=2)
```

Looking at missingness by variable and combinations of missingness using "aggr" from VIM package.  
```{r}
vim_plot = aggr(train, numbers = TRUE, prop = c(TRUE, FALSE),cex.axis=.42)
#the cex.axis reduces size of text on x-axis so labels fit better
```
Delete columns with high number (greater than 2000) of missing values column  
```{r}
#train = train %>% select(-measurement_10, -measurement_11, -measurement_12, -measurement_13, -measurement_14, -measurement_15, -measurement_16, -measurement_17)
```

Only missingness is in TotalCharges. Very small percentage. Row-wise deletion should be fine.  
```{r}
#train = train %>% drop_na()
```


Visualization of Attributes and Failure Train 
```{r}
p1 = ggplot(train, aes(x = attribute_0, fill = failure)) + geom_bar(position = "fill") 
p2 = ggplot(train, aes(x = attribute_1, fill = failure)) + geom_bar(position = "fill")
p3 = ggplot(train, aes(x = attribute_2, fill = failure)) + geom_bar(position = "fill")
p4 = ggplot(train, aes(x = attribute_3, fill = failure)) + geom_bar(position = "fill")
grid.arrange(p1,p2,p3,p4)
```
```{r}
p1 = ggplot(train, aes(x = failure, y = attribute_0)) + geom_boxplot()
p2 = ggplot(train, aes(x = failure, y = attribute_1)) + geom_boxplot()
p3 = ggplot(train, aes(x = failure, y = attribute_2)) + geom_boxplot()
p4 = ggplot(train, aes(x = failure, y = attribute_3)) + geom_boxplot()
grid.arrange(p1,p2,p3,p4, ncol = 2)
```

Visualization of Measurement and Failure
```{r}
p1 = ggplot(train, aes(x = measurement_0, fill = failure)) + geom_bar(position = "fill")
p2 = ggplot(train, aes(x = measurement_1, fill = failure)) + geom_bar(position = "fill")
p3 = ggplot(train, aes(x = measurement_2, fill = failure)) + geom_bar(position = "fill")
p4 = ggplot(train, aes(x = measurement_3, fill = failure)) + geom_bar(position = "fill")
grid.arrange(p1,p2,p3,p4)
```



Analyzing TEST file
Load data from the test.csv file.  
```{r}
test = read_csv("test.csv")
```

Structure and summary Test
```{r}
str(test)
summary(test)
```

Factor conversion attributes. Convert and recode the attribute variable.
```{r}
test = test %>% mutate(attribute_0 = as_factor(attribute_0)) %>% 
              mutate(attribute_0 = fct_recode(attribute_0, "7" = "material_7", "5" = "material_5")) 
test = test %>% mutate(attribute_1 = as_factor(attribute_1)) %>% 
              mutate(attribute_1 = fct_recode(attribute_1, "8" = "material_8", "5" = "material_5", "6" = "material_6")) 
```

Convert all character variables to factors  
```{r}
test = test %>% mutate_if(is.character,as_factor)
str(test)
summary(test)
```

```{r}
#use "ggcorr" to create a correlation matrix with labels and correlation reported to two decimals
ggcorr(test, method=c("pairwise"),label = TRUE,label_color = "blue", label_round = 2, label_size = 2, hjust = 1, layout.exp=2)
```

```{r}
skim(test)
```

